RILS-ROLS: Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-10-30 doc/paper.tex
Som eplots added

The rate of success of \textsc{Rils}--\textsc{Rols} algorithm seems to slightly increase as the expected number of variables gets larger in case of the data with no noise;
	the increase is significant in the case of a higher noise in data. For example, the lowest percentages of success for \textsc{Rils}-\textsc{Rols} algorithm are noticed for noisy data where exact models consist of just one or two variables. It seems that in these cases the algorithm is more likely favorizing including additional variables at the cost of increasing model accuracy  in order to improve the fitness value over iterations. In the presence of zero level od noise, OLS method will find much easier a smaller model (on $\approx$ 45\% cases) of a high  accuracy w.r.t.\ the input data. In the next iterations, the method will hardly intend for adding variables in the current best solution at the cost of improving fitness value, as the size of model will contribute more to its increase as the first two product terms in~\ref{eq:fitness} are already small (close to 1).     
	%Obtaining an accurate model with a few variables in the presence of a high level of nose in the data

\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=$variable\;count\;in\;exact\;models$,
			ylabel=$exact\;percentage$,
			xmin=0, xmax=5,
			ymin=0, ymax=100,
			xtick={1,2,3,4},
			xticklabels={1,2,3,4},   % <---
			ytick={0,10,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,43.33)
				(2,43.33)
				(3,46.67)
				(4,46.67)
			};
			\addlegendentry{No noise}

@@

			\addplot[smooth,color=green,mark=o]
			plot coordinates {
				(1,3.33)
				(2,3.33)
				(3,3.33)
				(4,16.67)
			};
			\addlegendentry{Noise level 0.01}
		\end{axis}
	\end{tikzpicture}
	\captionof{figure}{Exact solution percentages for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
	\label{fig:compExact_noise_varcnt}
\end{center}

%\begin{center}
%	\begin{tikzpicture}
%		\begin{axis}[
%			xlabel=$variable\;count$,
%			ylabel=$exact\;percentage$,
%			xmin=0, xmax=5,
%			ymin=0, ymax=100,
%			xtick={1,2,3,4},
%			xticklabels={1,2,3,4},   % <---
%			ytick={0,10,...,100}
%			]
%			\addplot[smooth,mark=*,blue] plot coordinates {
%				(1,83.33)
%				(2,66.67)
%				(3,73.33)
%				(4,46.67)
%			};
%			\addlegendentry{No noise}
%			
%			\addplot[smooth,color=red,mark=x]
%			plot coordinates {
%				(1,56.67)
%				(2,63.33)
%				(3,66.67)
%				(4,46.67)
%			};
%			\addlegendentry{Noise level 0.001}
%			
%			\addplot[smooth,color=green,mark=o]
%			plot coordinates {
%				(1,23.33)
%				(2,43.33)
%				(3,53.33)
%				(4,26.67)
%			};
%			\addlegendentry{Noise level 0.01}
%		\end{axis}
%	\end{tikzpicture}
%	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
%	\label{fig:compR2_noise_varcnt}
%\end{center}

>>>>

The rate of success of \textsc{Rils}--\textsc{Rols} algorithm seems to slightly increase as the expected number of variables gets larger in case of the data with no noise;
	the increase is significant in the case of a higher noise in data. For example, the lowest percentages of success for \textsc{Rils}-\textsc{Rols} algorithm are noticed for noisy data where exact models consist of just one or two variables. It seems that in these cases the algorithm is more likely in favor of including additional variables at the cost of increasing model accuracy  in order to improve the fitness value over iterations. In the presence of zero level od noise, OLS method will find much easier a smaller model (on $\approx$ 45\% cases) of a high  accuracy w.r.t.\ the input data. In the next iterations, the method will hardly intend for adding variables in the current best solution at the cost of improving fitness value, as the size of model will contribute more to its increase as the first two product terms in~\ref{eq:fitness} are already small (close to 1).      \fxnote{TODO: need to be further expanded...}
	%Obtaining an accurate model with a few variables in the presence of a high level of nose in the data

\begin{figure}[!ht]

	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numvars_vs_exact_correct_no_noise.eps}
		\caption{No noise}
		\label{fig:noNoise}
	\end{subfigure}
		\hfill
		\begin{subfigure}[b]{0.4\textwidth}

			\includegraphics[  width=180pt,height=140pt]{plots/numvars_vs_exact_correct_noise0_001.eps}
			\caption{Level of noise equal to 0.001}
			\label{fig:noise0.001}
		\end{subfigure}
		\centering
		\begin{subfigure}[b]{0.40\textwidth}
					\includegraphics[  width=180pt,height=140pt]{plots/numvars_vs_exact_correct_noise0_01.eps}
					\caption{Level of noise equal to 0.01.}
					\label{fig:noise0.01}
				\end{subfigure}
		\caption{Exact solution percentages for varying levels of noise and variable counts.}
		\label{fig:compExact_noise_varcnt}
	\end{figure}

\begin{figure}[!ht]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_exact_correct_no_noise.eps}
		\caption{No noise}
		\label{fig:size-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_exact_correct_noise0_001.eps}
		\caption{Level of noise equal to 0.001}
		\label{fig:size-noise0.001}
	\end{subfigure}
	\centering
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_exact_correct_noise0_01.eps}
		\caption{Level of noise equal to 0.01.}
		\label{fig:size-noise0.01}
	\end{subfigure}
	\caption{Exact solution percentages for varying levels of noise and exact model sizes.}
	\label{fig:compExact_noise_varsize}
\end{figure}

\begin{figure}[H]
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/numsize_vs_time_no_noise.eps}
		\caption{No noise}
		\label{fig:time-no-noise}
	\end{subfigure}
	\hfill
	\begin{subfigure}[b]{0.4\textwidth}

		\includegraphics[  width=180pt,height=140pt]{plots/vars_vs_time_noise0_001.eps}
		\caption{Level of noise equal to 0.001}
		\label{fig:time-noise0.001}
	\end{subfigure}
	\centering
	\begin{subfigure}[b]{0.40\textwidth}
		\includegraphics[  width=180pt,height=140pt]{plots/vars_vs_time_noise0_001.eps}
		\caption{Level of noise equal to 0.01.}
		\label{fig:time-noise0.01}
	\end{subfigure}
	\caption{Average runtime comparisons for varying levels of noise and variable counts.}
	\label{fig:runtime_rils_rols}
\end{figure}

 \begin{comment}
\begin{center}
	\begin{tikzpicture}
		\begin{axis}[
			xlabel=$variable\;count\;in\;exact\;models$,
			ylabel=$exact\;percentage$,
			xmin=0, xmax=5,
			ymin=0, ymax=100,
			xtick={1,2,3,4},
			xticklabels={1,2,3,4},   % <---
			ytick={0,10,...,100}
			]
			\addplot[smooth,mark=*,blue] plot coordinates {
				(1,43.33)
				(2,43.33)
				(3,46.67)
				(4,46.67)
			};
			\addlegendentry{No noise}

@@

			\addplot[smooth,color=green,mark=o]
			plot coordinates {
				(1,3.33)
				(2,3.33)
				(3,3.33)
				(4,16.67)
			};
			\addlegendentry{Noise level 0.01}
		\end{axis}
	\end{tikzpicture}
	\captionof{figure}{Exact solution percentages for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
	\label{fig:compExact_noise_varcnt}
\end{center}
\end{comment}
%\begin{center}
%	\begin{tikzpicture}
%		\begin{axis}[
%			xlabel=$variable\;count$,
%			ylabel=$exact\;percentage$,
%			xmin=0, xmax=5,
%			ymin=0, ymax=100,
%			xtick={1,2,3,4},
%			xticklabels={1,2,3,4},   % <---
%			ytick={0,10,...,100}
%			]
%			\addplot[smooth,mark=*,blue] plot coordinates {
%				(1,83.33)
%				(2,66.67)
%				(3,73.33)
%				(4,46.67)
%			};
%			\addlegendentry{No noise}
%			
%			\addplot[smooth,color=red,mark=x]
%			plot coordinates {
%				(1,56.67)
%				(2,63.33)
%				(3,66.67)
%				(4,46.67)
%			};
%			\addlegendentry{Noise level 0.001}
%			
%			\addplot[smooth,color=green,mark=o]
%			plot coordinates {
%				(1,23.33)
%				(2,43.33)
%				(3,53.33)
%				(4,26.67)
%			};
%			\addlegendentry{Noise level 0.01}
%		\end{axis}
%	\end{tikzpicture}
%	\captionof{figure}{Percentages of solutions having $R^2 > 0.999$ for varying levels of noise and variable counts (only sizes ranging from 7 to 12)}
%	\label{fig:compR2_noise_varcnt}
%\end{center}
